```{r, include=FALSE}
# At the top of each .Rmd file have a preamble chunk to load anything you want to load secretly in the background. e.g.
library(inlabru)
library(ggplot2)
data(gorillas)

# I think we want each .Rmd file to be able to work as a stand-alone file
# so we don't need to compile the whole book or run previous chapters to
# check things are working here.
```


# Introduction

This is an introduction to these notes.

This is about inlabru.  inlabru has data on gorilla nests:

```{r 01_gorillas_plot, echo=FALSE}
# echo=FALSE hides the code but it is still run
ggplot() +
  gg(gorillas$nests) + 
  gg(gorillas$boundary) + 
  coord_equal()
```

## Useful Resources

This section includes links to useful websites, tutorials, books, and papers.

* [INLA website](https://www.r-inla.org/)
* [INLA Discussion Group](https://groups.google.com/g/r-inla-discussion-group)
* [Tutorials for inlabru version 2.1.13](https://sites.google.com/inlabru.org/inlabru/tutorials)
* [Tutorials for inlabru version 2.2.4 and newer](https://inlabru-org.github.io/inlabru/)
* [Online Course Topics on INLA, Bayesian modeling, the SPDE approach, and spatial/temporal modeling - Haakon Bakka](https://haakonbakkagit.github.io/index.html)
* [Vignettes on Spatial Data Structures and Building a Mesh in INLA - Jafet Belmont](https://rpubs.com/jafet089)
* [Helper functions for mesh construction with the barrier model - James Grecian](https://github.com/jamesgrecian/mapr)
* [R-INLA Tutorial on SPDE Models - Elias T. Krainski, Finn Lindgren, Daniel Simpson and Håvard Rue](https://folk.ntnu.no/fuglstad/Lund2016/Session6/spde-tutorial.pdf)
* [Bayesian Inference with INLA - Virgilio Gómez-Rubio](https://becarioprecario.bitbucket.io/inla-gitbook/)

## Reading List

This section contains papers that may be useful to read for future sessions. 

* [How do covariates interact with spatial random effects?](https://www.tandfonline.com/doi/abs/10.1198/tast.2010.10052) (for the sessions on fields and covariates)
* [Choosing block spatial resolution in block cross-validation](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13107) (for the session on model assessment)

## Meeting Series Timeline
Date | Meeting Topic
---- | -------------
15/03/2022 | Planning discussion topics
22/03/2022 | Spatial data structures
29/03/2022 | Basics of mesh construction
05/04/2022 | <span style="color: red;">N/A - no meeting</span>
12/04/2022 | Catch-up meeting
19/04/2022 | How do inlabru model components work under the hood?
26/04/2022 | <span style="color: red;">N/A - no meeting</span>
03/05/2022 | <span style="color: red;">N/A - no meeting</span>
10/05/2022 | Inlabru model components and spatial data structures
17/05/2022 | <span style="color: red;">N/A - no meeting</span>
24/05/2022 | Boundaries
31/05/2022 | Mesh Properties
07/06/2022 | Prior choice
14/06/2022 | Covariates
21/06/2022 | Fields
28/06/2022 | <span style="color: red;">N/A - no meeting</span>
05/07/2022 | Reducing computational cost
12/07/2022 | Model comparison
19/07/2022 | Model assessment
26/07/2022 | Presenting and interpreting results
02/08/2022 | Joint models
09/08/2022 | Sampling
16/08/2022 | Excursions